---
title: "Time series plots"
author: "Abby Lewis and Adrienne Breef-Pilz"
date: "2022-09-08 edit 2023-06-16"
output: html_document
---

This file creates three summary time series plots that are in the main manuscript:
- Figure OXYGEN
- Figure BLOOM EFFECTS
- Figure PHYTOS

To do so, this file loads chemistry and fluora data from EDI, then loads saved values for metals, GHGs, thermocline depth, light attenuation, CTD data, and calculated fluora metrics, drawn from other scripts. 

TABLE OF CONTENTS:
Code chunk 1: Load packages, define drawdown dates
Code chunk 2: Load and format chemistry data
Code chunk 3: Load fluora data from EDI
Code chunk 4: Load other saved data for line plots
Code chunk 5: Compile Stats for Oxygen Figure
Code chunk 6: Make line plot of metals data, DO, and thermocline depth (Figure OXYGEN)
Code chunk 7: Compile stats for Bloom Effects Figure
Code chunk 8: Make line plot of nutrients, phytos, DOC, and CO2 (Figure BLOOM EFFECTS)
Code chunk 9: Compile stats for Phytos Figure
Code chunk 10: Make line plot of turbidity, euphotic zone, peak width, and peak depth (Figure PHYTOS)


Code chunk 1: Load packages, define drawdown dates
```{r Load packages and define drawdown dates, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(akima)
library(colorRamps)
library(ggh4x)
library(cowplot)

#Define drawdown lines
lines = data.frame(Dates = as.Date(c("2022-06-28", "2022-05-19","2022-05-28", "2022-07-01")), desc = c("closed","opened", "decreasing", "stable"), type = c("close","open","open","close"), status = c("init","init","final","final"), Year = 2022, year_class = "2022")
#Only plot pipe manipulations for now
lines = lines%>%
  filter(status == "init")
#Re-format for rectangle instead of discrete lines
lines_rect = lines%>%
  select(-type)%>%
  pivot_wider(names_from = desc, values_from = Dates)
```

Code chunk 2: Load and format chemistry data
```{r Load and format chemistry data}
inUrl1  <- "https://pasta.lternet.edu/package/data/eml/edi/199/11/509f39850b6f95628d10889d66885b76" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
 dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
        , col.names=c(
                    "Reservoir",     
                    "Site",     
                    "DateTime",     
                    "Depth_m",     
                    "Rep",     
                    "TN_ugL",     
                    "TP_ugL",     
                    "NH4_ugL",     
                    "NO3NO2_ugL",     
                    "SRP_ugL",     
                    "DOC_mgL",     
                    "DIC_mgL",     
                    "DC_mgL",     
                    "DN_mgL",     
                    "Flag_DateTime",     
                    "Flag_TN_ugL",     
                    "Flag_TP_ugL",     
                    "Flag_NH4_ugL",     
                    "Flag_NO3NO2_ugL",     
                    "Flag_SRP_ugL",     
                    "Flag_DOC_mgL",     
                    "Flag_DIC_mgL",     
                    "Flag_DC_mgL",     
                    "Flag_DN_mgL"    ), check.names=TRUE)
               
unlink(infile1)

#QAQC following conversation with CCC and HLW
dt1 = dt1%>%
  filter(hour(DateTime)>8,hour(DateTime)<17)%>%
  mutate(DateTime=as.Date(DateTime))%>%
  rename(PO4_ugL = SRP_ugL)%>%
  mutate(PO4_ugL = ifelse(PO4_ugL>TP_ugL,NA,PO4_ugL),
         TP_ugL = ifelse(Flag_TP_ugL%in%c(5,9)|TP_ugL>60,NA,TP_ugL),
         TN_ugL = ifelse(Flag_TN_ugL%in%c(5,9),NA,TN_ugL))

chem_full = dt1%>%
  filter(Reservoir=="BVR",
         Site==50)%>%
  filter(year(DateTime)%in%c(2021,2022))
#Add a "day of year" column with the year 2022 assigned
chem_full$Date_22<-chem_full$DateTime
chem_full$Year = year(chem_full$DateTime)
year(chem_full$Date_22)<-2022
```

Code chunk 3: Load fluora data from EDI
```{r Load fluora data from EDI}
#Load data
inUrl1  <- "https://pasta.lternet.edu/package/data/eml/edi/272/7/001cb516ad3e8cbabe1fdcf6826a0a45"
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
 dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
        , col.names=c(
                    "Reservoir",     
                    "Site",     
                    "DateTime",     
                    "Depth_m",     
                    "GreenAlgae_ugL",     
                    "Bluegreens_ugL",     
                    "BrownAlgae_ugL",     
                    "MixedAlgae_ugL",     
                    "YellowSubstances_ugL",     
                    "TotalConc_ugL",     
                    "Transmission",     
                    "Temp_degC",     
                    "RFU_525nm",     
                    "RFU_570nm",     
                    "RFU_610nm",     
                    "RFU_370nm",     
                    "RFU_590nm",     
                    "RFU_470nm",     
                    "Flag_GreenAlgae",     
                    "Flag_BluegreenAlgae",     
                    "Flag_BrownAlgae",     
                    "Flag_MixedAlgae",     
                    "Flag_TotalConc",     
                    "Flag_Temp",     
                    "Flag_Transmission",     
                    "Flag_525nm",     
                    "Flag_570nm",     
                    "Flag_610nm",     
                    "Flag_370nm",     
                    "Flag_590nm",     
                    "Flag_470nm"    ), check.names=TRUE)
               
unlink(infile1)

#Extract surface data
fluora_surface=dt1%>%
  filter(Reservoir=="BVR",
         Site==50,
         Depth_m<1)%>%
  mutate(DateTime = as.Date(DateTime))%>%
  group_by(DateTime)%>%
  mutate(min_depth = min(Depth_m))%>%
  filter(Depth_m==min_depth)%>%
  filter(year(DateTime)%in%c(2021,2022))
```


Code chunk 4: Load other saved data for line plots
```{r Load other saved data for line plots}
thermo_depth = read.csv("./Data/thermocline_depth.csv")%>%
  filter(year(Date)%in%c(2022,2021))

light = read.csv("./Data/attenuation_calc.csv")%>%
  filter(year(Date)%in%c(2021,2022))

ctd = read.csv("./Data/CTD at 0.1m.csv")%>%
  filter(year(Date)%in%c(2021,2022))

ghg = read.csv("./Data/GHG at 0.1m.csv")%>%
  filter(year(DateTime)%in%c(2021,2022))

metals_raw = read.csv("./Data/Metals at 0.1m.csv")%>%
  filter(year(Date)%in%c(2021,2022))
#Additional QAQC needed (Days where soluble>total)
metals = metals_raw%>%
  mutate(TFe_mgL = ifelse(Date=="2022-05-09",metals_raw$SFe_mgL[metals_raw$Date=="2022-05-09"],TFe_mgL),
         TMn_mgL = ifelse(Date=="2022-05-09",metals_raw$SMn_mgL[metals_raw$Date=="2022-05-09"],TMn_mgL),
         SFe_mgL = ifelse(Date=="2022-05-09",metals_raw$TFe_mgL[metals_raw$Date=="2022-05-09"],SFe_mgL),
         SMn_mgL = ifelse(Date=="2022-05-09",metals_raw$TMn_mgL[metals_raw$Date=="2022-05-09"],SMn_mgL),
         TFe_mgL = ifelse(Date=="2022-06-07",NA,TFe_mgL),
         TMn_mgL = ifelse(Date=="2022-06-07",NA,TMn_mgL))

cmax_depth = read.csv("./Data/FP_CmaxDepth.csv")%>%
  filter(year(Date)%in%c(2021,2022))

peak_width = read.csv("./Data/FP_PeakWidth.csv")%>%
  filter(year(Date)%in%c(2021,2022))
```


Code chunk 5: Make line plot of metals data, DO, and thermocline depth (Figure OXYGEN)
```{r Stats for Oxygen figure}
#combine relevant data
sum_plot3 = thermo_depth%>%
  mutate(DateTime = as.Date(Date))%>%
  select(-Date)%>%
  full_join(ctd%>%mutate(DateTime = as.Date(Date))%>%select(-Date,-Year,-Depth_m))%>%
  full_join(metals%>%mutate(DateTime = as.Date(Date))%>%select(-Date,-Date_22))%>%
  select(DateTime, TFe_mgL,TMn_mgL,SFe_mgL,SMn_mgL,DOsat_percent,thermo)%>%
  pivot_longer(cols = c(TFe_mgL,TMn_mgL,SFe_mgL,SMn_mgL,DOsat_percent,thermo))%>% #Pivot for facet wrap
  filter(!is.na(value),
         month(DateTime)>4,
         month(DateTime)<9)%>%
  mutate(Year = year(DateTime),
         color = name,
         name = ifelse(name%in%c("SFe_mgL","TFe_mgL"),"Fe_mgL",name),
         name = ifelse(name%in%c("SMn_mgL","TMn_mgL"),"Mn_mgL",name))%>% #Re-format to put solubles in the same plot as totals
  mutate(name = factor(name, levels=c("thermo","DOsat_percent","Fe_mgL","Mn_mgL")),
         color = factor(color, levels=c("thermo","DOsat_percent","TFe_mgL","TMn_mgL","SFe_mgL","SMn_mgL")),
         lty = ifelse(color %in%c("SFe_mgL","SMn_mgL"),"dashed","solid"))

#Calculate maximum values
maxes = sum_plot3%>%
  group_by(name, color,lty,Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            min_date = DateTime[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL","DOsat_percent"),min_date,max_date))

maxes_sum = sum_plot3%>%
  group_by(name,color, Year)%>%
  summarize(max = max(value),
            max_date = DateTime[which.max(value)],
            min_date = DateTime[which.min(value)])%>%
  group_by(name,color)%>%
  mutate(dif = max[Year==2022]-max[Year==2021],
         pct = (max[Year==2022]-max[Year==2021])/max[Year==2021]*100,
         date_dif = max_date[Year==2022]-(max_date[Year==2021]+365))

mins_sum = sum_plot3%>%
  group_by(name,color, Year)%>%
  summarize(min = min(value),
            min_date = DateTime[which.min(value)],
            min_date = DateTime[which.min(value)])%>%
  group_by(name,color)%>%
  mutate(dif = min[Year==2022]-min[Year==2021],
         pct = (min[Year==2022]-min[Year==2021])/min[Year==2021]*100,
         date_dif = min_date[Year==2022]-(min_date[Year==2021]+365))

#Stats
sum_plot3%>%
  group_by(name, color,Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            max = max(value),
            min_date = DateTime[which.min(value)],
            min = min(value))%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL","DOsat_percent"),min_date,max_date))

# New facet label names for sum_plot
fac.labs <- c("Surface DO\n(mgL)","Fe (mg/L)","Mn (mg/L)","Thermocline\ndepth (m)","Surface DO\n(% saturation)")
names(fac.labs) <- c("DO_mgL","Fe_mgL","Mn_mgL","thermo","DOsat_percent")
scales <- list(
  scale_y_reverse(limits = c(NA,0)),
  scale_y_continuous(),
  scale_y_continuous(),
  scale_y_continuous()
)
```


```{r Plot for oxygen Figure}
a<-sum_plot3%>%
  filter(name=="thermo")%>%
  mutate(show_points = ifelse(name == "thermo","n","y"))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  #geom_point(aes(alpha = show_points, color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, lty = lty, color = name))+
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19177, color = "thermo"), lwd = 2, alpha = 0.3, show.legend = F)+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  scale_linetype_manual(values = c("solid","11"),
                        breaks = c("solid","dashed"),
                        labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("darkblue","cyan4","lightsalmon","coral4"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(y="Thermocline \ndepth (m)",
       x="")+
  theme(strip.placement = "outside", 
        strip.text.x = element_text(
        size = 12, face = "bold"
        ),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        #legend.title=element_blank(),
        #legend.background = element_blank(),
        #legend.key= element_blank(),
        #legend.spacing.y = unit(0, 'cm'),
        #legend.text=element_text(size=7),
        plot.margin = unit(c(0.1,0,0,0), "cm")
        )+
  #guides(lty = "none",
         #color = guide_legend(byrow = TRUE,keyheight = 0.75),
         #alpha = "none")+
   facetted_pos_scales(y = scales)+
  force_panelsizes(rows = unit(.8, "in"),
                   cols = unit(2.5, "in"))


b<-sum_plot3%>%
  filter(name=="DOsat_percent")%>%
  mutate(show_points = ifelse(name == "thermo","n","y"))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, lty = lty, color = name))+
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19184, color = "DOsat_percent"), lwd = 2, alpha = 0.3, show.legend = F)+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  scale_linetype_manual(values = c("solid","11"),
                        breaks = c("solid","dashed"),
                        labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("cyan4","lightsalmon","coral4"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(y="  Surface DO \n (% saturation)",
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-.88,0,-.9,0), "cm")
        )+
  #guides(lty = "none",
         #color = guide_legend(byrow = TRUE,keyheight = 0.75),
         #alpha = "none")+
   #facetted_pos_scales(y = scales)+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Filter Iron out for the plot
fe<-sum_plot3%>%
  filter(name=="Fe_mgL")

# combine color and linetype in a new column 
fe$comb <- paste(fe$color, fe$lty)

# rename the column for the legend
fe<-fe%>%
  mutate(comb2=ifelse(comb=="SFe_mgL dashed","Soluble Fe", "Total Fe"))

fe$comb2 <- factor(fe$comb2, levels = c("Total Fe", "Soluble Fe"))

c<-fe%>%
ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
 geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19184, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19178, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  #scale_linetype_manual(values = c("solid","11"),
                        #breaks = c("solid","dashed"),
                        #labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("lightsalmon", "lightsalmon","lightsalmon"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(y="Fe (mg/L)",
       x="")+
      #color  = "Guide name", 
      #linetype = "Guide name")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.18,0.66),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-0.9,0,0,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol=2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
   #facetted_pos_scales(y = scales)+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))


# Filter Mn out for the plot
mn<-sum_plot3%>%
  filter(name=="Mn_mgL")

# combine color and linetype in a new column 
mn$comb <- paste(mn$color, mn$lty)

# rename the column for the legend
mn<-mn%>%
  mutate(comb2=ifelse(comb=="SMn_mgL dashed","Soluble Mn", "Total Mn"))

# Reorder so totals is a solid line
mn$comb2 <- factor(mn$comb2, levels = c("Total Mn", "Soluble Mn"))

d<-mn%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19184, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19170, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  #scale_linetype_manual(values = c("solid","11"),
                        #breaks = c("solid","dashed"),
                        #labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("coral4","coral4","coral4"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(y="Mn (mg/L)",
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        #axis.text.x = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.18,0.66),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.45,0,0,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
   #facetted_pos_scales(y = scales)+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

bio<-plot_grid(a, b,c,d, nrow = 4, align = "v")


 ggsave(file="./Figs/Abby_money_plot_biogeo_wLegend.jpeg",bio, dpi = 300, width = 6.5, height = 4, bg = "white")
 
```

Code chunk 6: Make line plot of nutrients, phytos, DOC, and CO2 (Figure BLOOM EFFECTS)

```{r bloom effects stats}

#Combine all data
sum_plot4 = chem_full%>%
  filter(Depth_m ==0.1)%>%
  select(DateTime, TP_ugL, TN_ugL,PO4_ugL,DOC_mgL,DIC_mgL,NO3NO2_ugL,NH4_ugL, NO3NO2_ugL)%>%
  full_join(fluora_surface%>%
              filter(year(DateTime)%in% c(2021,2022)))%>%
  full_join(ghg%>%mutate(DateTime = as.Date(DateTime)))%>%
  mutate(TNTP = TN_ugL/TP_ugL)%>%
  select(DateTime, TP_ugL, TN_ugL,PO4_ugL,DOC_mgL,TotalConc_ugL,NH4_ugL,co2_umolL,Bluegreens_ugL)%>%
  pivot_longer(cols = c(TP_ugL, TN_ugL,PO4_ugL,DOC_mgL,TotalConc_ugL,co2_umolL,NH4_ugL,Bluegreens_ugL))%>% #pivot for facet wrap
  filter(!is.na(value),
         month(DateTime)>4,
         month(DateTime)<9)%>%
  mutate(Year = year(DateTime),
         color = name,
         name = ifelse(name%in%c("TotalConc_ugL","Bluegreens_ugL"),"Phyto",name), #Re-format to get cyanos on the same plot as totals
         name = ifelse(name%in%c("TP_ugL","PO4_ugL"),"Phosphorus",name), #Re-format to get solubles on the same plot as totals
         name = ifelse(name%in%c("TN_ugL","NO3NO2_ugL","NH4_ugL"),"Nitrogen",name),
         lty = ifelse(color%in%c("NO3NO2_ugL","NH4_ugL","PO4_ugL","Bluegreens_ugL"),"dashed","solid"))%>%
  mutate(name = factor(name, levels=c("Phosphorus","Nitrogen","Phyto","DOC_mgL","co2_umolL")),
         color = factor(color, levels=c("TP_ugL","PO4_ugL","TN_ugL","NH4_ugL","TotalConc_ugL","Bluegreens_ugL","DOC_mgL","co2_umolL"),
                        labels=c("TP","PO4","TN","NH4","Total phytoplankton", "Cyanobacteria", "DOC", "CO2"))) #Re-label colors

#Calculate maxes
maxes = sum_plot4%>%
  group_by(name,color, Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            min_date = DateTime[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL"),min_date,max_date))

maxes_sum = sum_plot4%>%
  group_by(name,color, Year)%>%
  summarize(max = max(value))%>%
  group_by(name,color)%>%
  mutate(dif = max[Year==2022]-max[Year==2021],
         pct = (max[Year==2022]-max[Year==2021])/max[Year==2021]*100)

#Stats
maxes%>%
  mutate(max_date=as.Date(max_date, origin="1970-01-01"))
as.Date("2022-07-05")-as.Date("2022-06-27")
as.Date("2022-07-11")-as.Date("2022-06-27")
as.Date("2022-07-18")-as.Date("2022-06-27")
as.numeric(maxes$max_date)

# New facet label names for sum_plot
fac.labs <- c("Phosphorus\n(µg/L)", "Nitrogen\n(µg/L)","Phytoplankton\n(µg/L)", "DOC\n(mg/L)", "CO2\n(µmol/L)", "NH4 (µg/L)")
names(fac.labs) <- c("Phosphorus", "Nitrogen","Phyto", "DOC_mgL", "co2_umolL", "NH4_ugL")
#Specify colors
my_colors = c("black","black","magenta4","magenta4","darkgreen","darkgreen","#8E4412","blue")
my_lines = c("solid","11","solid","11","solid","11","solid","solid")
labels=c("TP",bquote(PO[4]),"TN",bquote(NH[4]),"Total phytoplankton","Cyanobacteria","DOC",bquote(CO[2]))
names_orig=c("TP","PO4","TN","NH4","Total phytoplankton","Cyanobacteria","DOC","CO2")
#names(my_lines)=c("TP",expression(PO[4]),"Total phytoplankton","Cyanobacteria","DOC","CO2","NH4","TN")
```

```{r plots bloom effect}

#PLOT
# plot each one individual to plots and then merge them at the end

ph<-sum_plot4%>%
  filter(name=="Phosphorus")

# combine color and linetype in a new column 
ph$comb <- paste(ph$color, ph$lty)

# rename the column for the legend
ph<-ph%>%
  mutate(comb2=ifelse(comb=="TP solid","Total P", "Phosphate"))

ph$comb2 <- factor(ph$comb2, levels = c("Total P", "Phosphate"))


aa<-ph%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  geom_vline(aes(xintercept = 19170, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19163, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #scale_linetype_manual(values = my_lines,
                        #labels = labels,
                        #breaks = names_orig
                        #)+
  scale_color_manual(values = c("black","black","black"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(#y=expression(paste("Phosphorus \n ","(",mu, "g/L)")),
       y=expression(atop("Phosphorus", paste("(",mu, "g/L)"))),
       x="")+
  theme(strip.placement = "outside", 
        strip.text.x = element_text(
        size = 12, face = "bold"
        ),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        #legend.position = "none",
        legend.position = c(0.12,0.38),
        legend.justification = c(0,0),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(0.1,0,-1.2,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots Nitrogen
ni<-sum_plot4%>%
  filter(name=="Nitrogen")

ni$comb <- paste(ni$color, ni$lty)

# rename the column for the legend
ni<-ni%>%
  mutate(comb2=ifelse(comb=="TN solid","Total N", "Ammonium"))

ni$comb2 <- factor(ni$comb2, levels = c("Total N", "Ammonium"))


bb<-ni%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  geom_vline(aes(xintercept = 19170, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19191, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #scale_linetype_manual(values = my_lines,
                        #labels = labels,
                        #breaks = names_orig
                        #)+
  scale_color_manual(values = c("magenta4","magenta4","magenta4"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(#y=expression(paste("Nitrogen \n","(",mu, "g/L)")),
    y=expression(atop("Nitrogen", paste("(",mu, "g/L)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.12,0.38),
        legend.justification = c(0,0),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.2,0,-2.3,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots Phytoplankton
pl<-sum_plot4%>%
  filter(name=="Phyto")

pl$comb <- paste(pl$color, pl$lty)

# rename the column for the legend
pl<-pl%>%
  mutate(comb2=ifelse(comb=="Total phytoplankton solid","Total phytoplankton", "Cyanobacteria"))

pl$comb2 <- factor(pl$comb2, levels = c("Total phytoplankton", "Cyanobacteria"))

cc<-pl%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  geom_vline(aes(xintercept = 19170, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19170, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #scale_linetype_manual(values = my_lines,
                        #labels = labels,
                        #breaks = names_orig
                        #)+
  scale_color_manual(values = c("darkgreen","darkgreen","darkgreen"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(#y=expression(paste("Nitrogen \n","(",mu, "g/L)")),
    y=expression(atop("Phytoplankton", paste("(",mu, "g/L)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.12,0.38),
        legend.justification = c(0,0),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-2,0,-2,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots DOC

dd<-sum_plot4%>%
  filter(name=="DOC_mgL")%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, color = name))+
  geom_vline(aes(xintercept = 19178, color = "DOC"), lwd = 2, alpha = 0.3, show.legend = F)+
  scale_linetype_manual(values = my_lines,
                        labels = labels,
                        breaks = names_orig
                        )+
  scale_color_manual(values = c("#8E4412","#8E4412"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(#y="DOC \n (mg/L)",
    y=expression(atop("DOC", paste("(mg/L)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.4,0,-0.4,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots CO2

ee<-sum_plot4%>%
  filter(name=="co2_umolL")%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, color = name))+
  geom_vline(aes(xintercept = 19184, color = "CO2"), lwd = 2, alpha = 0.3, show.legend = F)+
  scale_linetype_manual(values = my_lines,
                        labels = labels,
                        breaks = names_orig
                        )+
  scale_color_manual(values = c("blue","blue"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(
  y=expression(atop("CO"[2], paste("(", mu, "mol/L)"))),
    #y=expression(atop("DOC", paste("(",mu, "g/L)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        #axis.text.x = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.7,0,0,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

bloom<-plot_grid(aa,bb,cc,dd,ee, nrow = 5, align = "v", hjust = -0.5)

ggsave("./Figs/Abby_money_plot_doc_short_wlegend.jpeg",bloom, dpi = 300, width = 6.5, height = 5.5, bg = "white")
```

TNTP
```{r}
#Combine all data
tntp = chem_full%>%
  filter(Depth_m ==0.1)%>%
  select(DateTime, TP_ugL, TN_ugL)%>%
  full_join(fluora_surface%>%
              filter(year(DateTime)%in% c(2021,2022)))%>%
  mutate(TNTP = TN_ugL/TP_ugL)%>%
  select(DateTime, TNTP, TotalConc_ugL, Bluegreens_ugL)%>%
  pivot_longer(cols = c(TNTP, TotalConc_ugL, Bluegreens_ugL))%>% #pivot for facet wrap
  filter(!is.na(value),
         month(DateTime)>4,
         month(DateTime)<9)%>%
  mutate(Year = year(DateTime),
         color = name,
         name = ifelse(name%in%c("TotalConc_ugL","Bluegreens_ugL"),"Phyto",name), #Re-format to get cyanos on the same plot as totals
         lty = ifelse(color%in%c("Bluegreens_ugL"),"dashed","solid"))%>%
  mutate(name = factor(name, levels=c("TNTP","Phyto")),
         color = factor(color, levels=c("TotalConc_ugL","Bluegreens_ugL","TNTP"),
                        labels=c("Total phytoplankton", "Cyanobacteria","TNTP"))) #Re-label colors

#Calculate maxes
maxes = tntp%>%
  group_by(name,color, Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            min_date = DateTime[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL"),min_date,max_date))

maxes_sum = tntp%>%
  group_by(name,color, Year)%>%
  summarize(max = max(value))%>%
  group_by(name,color)%>%
  mutate(dif = max[Year==2022]-max[Year==2021],
         pct = (max[Year==2022]-max[Year==2021])/max[Year==2021]*100)

#Stats
maxes%>%
  mutate(max_date=as.Date(max_date, origin="1970-01-01"))
as.numeric(maxes$max_date)

# New facet label names for sum_plot
fac.labs <- c("Phytoplankton\n(µg/L)", "TN:TP")
names(fac.labs) <- c("Phyto", "TNTP")
#Specify colors
my_colors = c("darkgreen","darkgreen","black")
my_lines = c("solid","11","solid")

tntp%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(y = value, x = DateTime,color = color))+
  geom_line(aes(y = value, lty=color,x = DateTime,color = color))+
  geom_vline(aes(xintercept = max_date, color = color), data = maxes, lwd = 2, alpha = 0.5, show.legend = F)+
  #geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  scale_linetype_manual(values = c("solid","11","solid"),
                        )+
  scale_color_manual(values = my_colors, 
                     )+
  facet_grid(rows = vars(name), cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  theme(strip.placement = "outside", 
        strip.text.x = element_text(
        size = 12, face = "bold"
        ),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        legend.position = "bottom",
        legend.title = element_blank())+
  guides(color = guide_legend(byrow = TRUE,keyheight = 0.75))+
  ylab("")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))
ggsave("./Figs/TNTP.jpeg", dpi = 300, width = 6.5, height = 3)

before = tntp$value[tntp$DateTime=="2022-05-23"&tntp$name=="TNTP"]
after = tntp$value[tntp$DateTime=="2022-06-27"&tntp$name=="TNTP"]
(after-before)/before*100
```

Code chunk 7: Make line plot of turbidity, euphotic zone, peak width, and peak depth (Figure PHYTOS)

```{r phytos stats}

sum_plot5 = fluora_surface%>%
  full_join(light%>%mutate(DateTime = as.Date(Date))%>%select(-Date))%>%
  full_join(ctd%>%mutate(DateTime = as.Date(Date))%>%select(-Date,-Year,-Depth_m))%>%
  full_join(cmax_depth%>%mutate(DateTime = as.Date(Date))%>%select(-Date))%>%
  full_join(peak_width%>%mutate(DateTime = as.Date(Date))%>%select(-Date))%>%
  select(DateTime, Turbidity_NTU, Zeu, CmaxDepth_TotalConc_ugL, CmaxDepth_Bluegreens_ugL, PeakWidth_TotalConc_m, PeakWidth_Bluegreens_m)%>%
  pivot_longer(cols = c(Turbidity_NTU, Zeu, CmaxDepth_TotalConc_ugL, CmaxDepth_Bluegreens_ugL, PeakWidth_TotalConc_m, PeakWidth_Bluegreens_m))%>%
  filter(!is.na(value),
         month(DateTime)>4,
         month(DateTime)<9)%>%
  mutate(Year = year(DateTime),
         color = ifelse(name%in%c("CmaxDepth_TotalConc_ugL", "PeakWidth_TotalConc_m"),"TotalConc_ugL",
                        ifelse(name%in%c("PeakWidth_Bluegreens_m", "CmaxDepth_Bluegreens_ugL"),"Bluegreens_ugL",name)),
         name = ifelse(name%in%c("CmaxDepth_TotalConc_ugL", "CmaxDepth_Bluegreens_ugL"),"Cmax_depth",name),
         name = ifelse(name%in%c("PeakWidth_TotalConc_m", "PeakWidth_Bluegreens_m"),"PeakWidth",name),
         lty = ifelse(color == "Bluegreens_ugL","dashed","solid"))%>%
  mutate(name = factor(name, levels=c("Turbidity_NTU", "Zeu", "Cmax_depth", "PeakWidth")),
         color = factor(color, levels=c("Turbidity_NTU", "Zeu","TotalConc_ugL", "Bluegreens_ugL")))

#Calculate maxes
maxes = sum_plot5%>%
  group_by(name,color, Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            min_date = DateTime[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL","Cmax_depth", "PeakWidth"),min_date,max_date))

#Stats
sum_plot5%>%
  group_by(name,color, Year)%>%
  summarize(max_date = DateTime[which.max(value)],
            max = max(value),
            min_date = DateTime[which.min(value)],
            min = min(value))%>%
  mutate(max_date = ifelse(name %in% c("thermo","Zeu","DO_mgL","Cmax_depth", "PeakWidth"),min_date,max_date),
         max = ifelse(name %in% c("thermo","Zeu","DO_mgL","Cmax_depth", "PeakWidth"),min,max))%>%
  group_by(name,color)%>%
  mutate(pct = (max[Year==2022]-max[Year==2021])/max[Year==2021]*100)

# New facet label names for sum_plot
fac.labs <- c("Turbidity\n(NTU)", "Euphotic zone\ndepth (m)", "C[max]\ndepth (m)", "Peak width\n(m)")
names(fac.labs) <- c("Turbidity_NTU", "Zeu", "Cmax_depth", "PeakWidth")
my_colors = c("grey20","goldenrod","darkgreen","turquoise")
names(my_colors)=c("Turbidity_NTU", "Zeu","TotalConc_ugL", "Bluegreens_ugL")
scales <- list(
  scale_y_continuous(),
  scale_y_reverse(limits = c(NA,0)),
  scale_y_reverse(limits = c(NA,0)),
  scale_y_continuous()
)
```

```{r phytos with legend plot}

#PLOT
# plot each one individual to plots and then merge them at the end

a3<-sum_plot5%>%
  filter(name=="Turbidity_NTU")%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, color = name))+
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19178, color = "Turbidity_NTU"), lwd = 2, alpha = 0.3, show.legend = F)+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  scale_linetype_manual(values = c("solid","11"),
                        breaks = c("solid","dashed"),
                        labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("black","black"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  theme_bw()+
  labs(#y="Turbidity \n(NTU)",
       y=expression(atop("Turbidity", paste("(NTU)"))),
       x="")+
  theme(strip.placement = "outside", 
        strip.text.x = element_text(
        size = 12, face = "bold"
        ),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        #legend.title=element_blank(),
        #legend.background = element_blank(),
        #legend.key= element_blank(),
        #legend.spacing.y = unit(0, 'cm'),
        #legend.text=element_text(size=7),
        plot.margin = unit(c(0.1,0,-0.68,0), "cm")
        )+
  #guides(lty = "none",
         #color = guide_legend(byrow = TRUE,keyheight = 0.75),
         #alpha = "none")+
   facetted_pos_scales(y = scales)+
  force_panelsizes(rows = unit(.8, "in"),
                   cols = unit(2.5, "in"))

# Plots Euphotic Zone

b3<-sum_plot5%>%
  filter(name=="Zeu")%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(color = name,x = DateTime, y = value))+
  geom_line(aes(x = DateTime, y = value, color = name))+
  scale_alpha_manual(values = c(0,1))+
  geom_vline(aes(xintercept = 19178, color = "Turbidity_NTU"), lwd = 2, alpha = 0.3, show.legend = F)+
  #geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  scale_linetype_manual(values = c("solid","11"),
                        breaks = c("solid","dashed"),
                        labels = c("Total","Soluble"),)+
  scale_color_manual(values = c("goldenrod","goldenrod"))+
  facet_grid(rows = NULL,cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  scale_y_reverse()+
  theme_bw()+
  labs(#y=expression(paste("Nitrogen \n","(",mu, "g/L)")),
    #y="Euphotic zone \n depth (m)",
    y=expression(atop("Euphotic zone", paste("depth (m)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = "none",
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1,0,-1.6,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots Phytoplankton Cmax
cmax<-sum_plot5%>%
  filter(name=="Cmax_depth")

cmax$comb <- paste(cmax$color, cmax$lty)

# rename the column for the legend
cmax<-cmax%>%
  mutate(comb2=ifelse(comb=="TotalConc_ugL solid","Total phytoplankton", "Cyanobacteria"))

cmax$comb2 <- factor(cmax$comb2, levels = c("Total phytoplankton", "Cyanobacteria"))

c3<-cmax%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  geom_vline(aes(xintercept = 19163, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19178, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #scale_linetype_manual(values = my_lines,
                        #labels = labels,
                        #breaks = names_orig
                        #)+
  scale_color_manual(values = c("darkgreen","darkgreen"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  scale_y_reverse()+
  theme_bw()+
  labs(#y=expression(paste("Nitrogen \n","(",mu, "g/L)")),
    y=expression(atop("C"["max"], paste("depth (m)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.20,0.66),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.8,0,-1.4,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

# Plots for peak width

pkw<-sum_plot5%>%
  filter(name=="PeakWidth")

pkw$comb <- paste(pkw$color, pkw$lty)

# rename the column for the legend
pkw<-pkw%>%
  mutate(comb2=ifelse(comb=="TotalConc_ugL solid","Total phytoplankton", "Cyanobacteria"))

pkw$comb2 <- factor(pkw$comb2, levels = c("Total phytoplankton", "Cyanobacteria"))


d4<-pkw%>%
    ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x=DateTime,
                y=value,
                colour = comb2))+ 
  geom_line(aes(x=DateTime,
                y=value,
                colour = comb2,
                linetype = comb2)) +
  geom_vline(aes(xintercept = 19156, color = comb2), lwd = 2, alpha = 0.3, show.legend = F)+
  geom_vline(aes(xintercept = 19163, color = comb2), lwd = 2, alpha = 0.3, show.legend = F, linetype = "11")+
  #scale_linetype_manual(values = my_lines,
                        #labels = labels,
                        #breaks = names_orig
                        #)+
  scale_color_manual(values = c("darkgreen","darkgreen"))+
  facet_grid(rows = NULL, cols = vars(Year), scales = "free", switch = "y",
             labeller = labeller(name = fac.labs))+
  #scale_y_reverse()+
  theme_bw()+
  labs(#y=expression(paste("Nitrogen \n","(",mu, "g/L)")),
    y=expression(atop("Peak width", paste("(m)"))),
       x="")+
  theme(#strip.placement = "outside", 
        #strip.text.x = element_text(
        #size = 12, face = "bold"
        #),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        legend.position = c(0.30,0.77),
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7),
        plot.margin = unit(c(-1.7,0,-0.4,0), "cm")
        )+
  guides(lty = guide_legend(byrow = TRUE,keyheight = 0.75, ncol = 2),
         color = guide_legend(byrow = TRUE,keyheight = 0.75),
         alpha = "none")+
  force_panelsizes(rows = unit(.75, "in"),
                   cols = unit(2.5, "in"))

phytos<-plot_grid(a3,b3,c3,d4, nrow = 4, align = "v", hjust = -0.5)

ggsave("./Figs/Abby_money_plot_phytos_wlegend.jpeg",phytos, dpi = 300, width = 6.5, height = 4, bg = "white")
```