---
title: "CTD Processing - Drawdown"
author: "Abby Lewis"
date: "12/11/2019"
output: html_document
---

This file loads CTD and YSI data from EDI and generates three outputs: 
1) a dataframe with CTD data from 0.1 m depth (using YSI to fill data gaps)
2) a dataframe with processed CTD data in 0.1 m intervals (using YSI to fill data gaps)
3) a dataframe with calculated light attenuation outputs


Step 1: Load data and packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse)
library(lubridate)
library(rLakeAnalyzer)


# Load CTD data from EDI
ctd_edi <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/200/13/27ceda6bc7fdec2e7d79a6e4fe16ffdf")
#format date, filter to the reservoir and site we are using
ctd <- ctd_edi%>%
  mutate(Date = as.Date(DateTime, format = "%Y-%m-%d %H:%M:%S"))%>%
  dplyr::select(-DateTime)%>%
  filter(Reservoir=="BVR", 
         Site==50)


#Load YSI data from EDI
ysi_edi <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/198/9/b3bd353312f9e37ca392e2a5315cc9da")
#filter to the reservoir and site we are using
ysi <- ysi_edi%>%
  filter(Site == 50,
         Reservoir=="BVR")
```


Step 2: Format data and combine with YSI
```{r}
# Decrease CTD dataset size by trimming to 0.1m depth intervals
depths <- seq(0,11, by = .1)
newDepths <- depths
df.final.raw<- ctd %>% group_by(Date, Reservoir, Site) %>% slice(which.min(abs(as.numeric(Depth_m) - depths[1]))) #Create a new dataframe
df.final.raw$Depth_m <- newDepths[1]
for (i in 2:length(depths)){ #loop through all depths and add the closest values to the final dataframe
  ctd_atThisDepth <- ctd %>% group_by(Date, Reservoir, Site) %>% slice(which.min(abs(as.numeric(Depth_m) - depths[i])))
  ctd_atThisDepth <- ctd_atThisDepth%>%
    filter(abs(Depth_m-newDepths[i])<0.1) #only include if the measured depth is within 0.1 of the depth we are labeling it
  ctd_atThisDepth$Depth_m <- newDepths[i]
  df.final.raw <- rbind(df.final.raw,ctd_atThisDepth)
}


# Format YSI data
ysi_formatted <- ysi%>%
  mutate(Date = as.Date(DateTime))%>%
  select("Reservoir", "Site", "Date", "Depth_m", "DO_mgL", "DOSat","Temp_C")%>%
  #We are only using YSI to fill in gaps in CTD data. Select data for those gaps here
  filter(year(Date)==2021)%>%
  mutate(DO_mgL = ifelse(Date<=as.Date("2021-07-01")|Date>=as.Date("2021-12-01"),NA,DO_mgL),
         DOSat = ifelse(Date<=as.Date("2021-07-01")|Date>=as.Date("2021-12-01"),NA,DOSat),
         Temp_C = ifelse(Date<=as.Date("2021-07-01")|Date>=as.Date("2021-12-01"),NA,Temp_C))%>%
  mutate(method = "ysi")%>%
  filter(Date!="2021-05-31"|Depth_m<=10)


# Combine CTD and YSI data
df.final_export <- df.final.raw%>%
  mutate(Date = as.Date(Date),
         Year = year(Date),
         DO_mgL = ifelse(Date>as.Date("2021-07-01")&Date<as.Date("2021-12-01"),
                         NA,DO_mgL),
         DOsat_percent = ifelse(Date>as.Date("2021-07-01")&Date<as.Date("2021-12-01"),
                                NA,DOsat_percent))%>%
  full_join(ysi_formatted%>%dplyr::rename(DO_mgL_ysi = DO_mgL,
                                          DOsat_percent_ysi = DOSat,
                                          Temp_C_ysi = Temp_C))%>%
  group_by(Date, Site)%>%
  mutate(max = max(Depth_m),
         Temp_C = ifelse(is.na(Temp_C),Temp_C_ysi, Temp_C),
         DO_mgL=ifelse(is.na(DO_mgL),DO_mgL_ysi,DO_mgL),
         DOsat_percent=ifelse(is.na(DOsat_percent),DOsat_percent_ysi,DOsat_percent),
         Depth_m_sed = max-Depth_m,
         Year = year(Date))


# Prepare 0.1 m data for export
export <- df.final_export%>%
  filter(Reservoir=="BVR",
         Depth_m==0.1)
write.csv(export,"../Processed_data/CTD at 0.1m.csv",row.names = F)


# Export full ctd dataset
df.final <- df.final_export%>%
  filter(max>5) #Filter to full profiles
write.csv(df.final,"../Processed_data/CTD trimmed.csv",row.names = F)
```


Step 3: Calculate light attenuation/light extinction
```{r}
#Using the equation here
#https://www.esf.edu/efb/schulz/Limnology/Light.html

#Add PAR data from the YSI dataset
ctd_bvr <- ctd%>%
  full_join(ysi%>%select(DateTime,Depth_m,PAR_umolm2s)%>%filter(!is.na(PAR_umolm2s))%>%rename(Date = DateTime)%>%mutate(Date=as.Date(Date)))


#Calculate light attenuation
atten_k = ctd_bvr%>%
  filter(!is.na(PAR_umolm2s),
         !is.na(Depth_m))%>%
  group_by(Date)%>%
  filter(sum(Depth_m<0)>0)%>%
  mutate(I0 = mean(PAR_umolm2s[Depth_m<0], na.rm = T),
         PAR_umolm2s = ifelse(PAR_umolm2s==0,0.001,PAR_umolm2s)
         )%>%
  filter(Depth_m>0,
         !I0==0)%>%
  summarize(I0 = unique(I0),
            k = coef(lm(I(log(PAR_umolm2s)-log(I0))~ 0 + Depth_m)),
            r2 = summary(lm(I(log(PAR_umolm2s)-log(I0))~ 0 + Depth_m))$r.squared,
            Zeu = min(Depth_m[PAR_umolm2s<I0/100]),
            Zeu_0.1 = min(Depth_m[PAR_umolm2s<I0/1000]))%>%
  filter(r2>0.9)


#Export
write.csv(atten_k,"../Processed_data/attenuation_calc.csv",row.names = F)
```