---
title: "Physics"
author: "Abby Lewis"
date: "2022-08-30"
output: html_document
---

This file loads in-situ thermistor data and calculates lake physics metrics. Six output files are generated:
1) Full formatted thermistor data file
2) Schmidt stability
3) Maximum buoyancy frequency
4) Volume-weighted water temperature
5) Thermocline depth
6) Time series volume of each layer of the reservoir

Table of contents:
Step 1: Load packages
Step 2: Load and format thermistor data
Step 3: Format bathymetry data
Step 4: Calculate Schmidt stability
Step 5: Calculate maximum buoyancy frequency
Step 6: Calculate volume-weighted water temperature
Step 7: Calculate thermocline depths
Step 8: Calculate volume of each layer of the reservoir


Step 1: Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rLakeAnalyzer)
library(lubridate)
library(zoo)
source("../R/thermo.depth.density.R")
```


Step 2: Load and format thermistor data
```{r}
sensor <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/725/3/a9a7ff6fe8dc20f7a8f89447d4dc2038")
depth_offset <- read.csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.725.3&entityid=5927a50118644fa451badb3b84233bb7")

#Need the actual depth of the thermistor sensors, accounting for changing water level
bvr_new <- sensor%>%
  select(Reservoir,Site,DateTime,starts_with("ThermistorTemp"),Depth_m_13)%>%
  pivot_longer(cols = starts_with("ThermistorTemp"), 
               values_to = "Reading", 
               names_prefix = "ThermistorTemp_C_",
               names_to = "Position")%>%
  mutate(Position = as.numeric(Position))%>%
  full_join(depth_offset)

bvr_pre_05APR21 <- bvr_new%>%
  filter(DateTime<="2021-04-05 13:20")%>%
  mutate(Sensor_depth=Depth_m_13-Offset_before_05APR21)%>% #this gives you the depth of the thermistors from the surface
  mutate(Rounded_depth_hundreth=round(Sensor_depth, 2))#Round to the nearest hundredth 
  
bvr_post_05APR21=bvr_new%>%
  filter(DateTime>"2021-04-05 13:20")%>%
  mutate(Sensor_depth=Depth_m_13-Offset_after_05APR21)%>% #this gives you the depth of the thermistor from the surface
  mutate(Rounded_depth_hundreth=round(Sensor_depth, 2)) #Round to the nearest hundredth 

# combine the pre April 5th and the post April 5th. Drop if the sensor depth is NA because can't
# figure out the depth of the sensors. This will give you a depth for each sensor reading. 
cat_long=bvr_pre_05APR21%>%
  rbind(.,bvr_post_05APR21)%>%
  filter(!is.na(Sensor_depth),
         year(DateTime)%in% c(2021,2022))%>%
  select(-Offset_before_05APR21, -Offset_after_05APR21, -Distance_above_sediments)%>%
  mutate(Reservoir_depth=Depth_m_13+0.5)%>%
  select(Reservoir, Site,DateTime, Reading, Sensor_depth, Rounded_depth_hundreth, Depth_m_13, Reservoir_depth)

write.csv(cat_long,"../Processed_data/BVR_longoutput.csv", row.names = F)

#Format temperature data for lake analyzer
catwalk_format = cat_long%>%
  filter(!is.na(Reading))%>%
  mutate(Date = as.Date(as.POSIXct(DateTime)),
         Sensor_depth = round(Sensor_depth,1)
         )%>%
  group_by(Date)%>%
  mutate(Reservoir_depth = mean(Reservoir_depth))%>%
  group_by(Date, Sensor_depth)%>%
  dplyr::summarize(Reading=mean(Reading,na.rm = T),
                   Reservoir_depth = unique(Reservoir_depth))
```


Step 3: Format bathymetry data
```{r}
#Load bathymetry
bathy_raw <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1254/1/f7fa2a06e1229ee75ea39eb586577184")

bathy = bathy_raw%>%
  filter(Reservoir == "BVR")

#Need to create changing bathymetry for each date, which I will do in several steps
#First, get the water levels at each date
cat_depths = cat_long%>%
  mutate(Date = as.Date(as.POSIXct(DateTime)))%>%
  filter(!is.na(Reading))%>%
  group_by(Date)%>%
  dplyr::summarize(WaterLevel = mean(Reservoir_depth))
#Then create a dataframe with bathymetry at each date
flexible_bathy = cat_depths%>%
  mutate(Reservoir = "BVR")%>%
  full_join(bathy)%>%
  group_by(Date)%>%
  mutate(Depth_m = Depth_m - (max(Depth_m)-unique(WaterLevel)) + 1)%>%
  filter(Depth_m>=0)
```


Step 4: Calculate Schmidt stability
```{r}
#Calculate schmidt stability each day
schmidts = c()
for(date in unique(catwalk_format$Date)){
  temps = catwalk_format%>%
    filter(Date == date)
  baths = flexible_bathy%>%
    filter(Date==date)
  schmidts = c(schmidts, schmidt.stability(wtr = temps$Reading, depths = temps$Sensor_depth, bthA = baths$SA_m2, bthD = baths$Depth_m, sal = rep(0,length(temps$Reading))))
}
schmidt_df = data.frame(Date = unique(catwalk_format$Date), Schmidt = schmidts)%>%
  mutate(Year = year(Date))
schmidt_df$Date_22 = schmidt_df$Date
year(schmidt_df$Date_22)=2022

write.csv(schmidt_df, "../Processed_data/Schmidt.csv", row.names = F)
```


Step 5: Calculate maximum buoyancy frequency
```{r}
#Calculate buoyancy frequency each day
buoyancy_raw = catwalk_format%>%
  filter(!is.na(Reading))%>%
  mutate(Depth_simple = floor(Sensor_depth))%>%
  group_by(Date,Depth_simple)%>%
  dplyr::summarize(Temp_C = mean(Reading),
            Depth_m = mean(Sensor_depth))%>%
  arrange(Date,Depth_m)%>%
  group_by(Date)%>%
  filter((Depth_m-lag(Depth_m)>.4)|(Depth_m-lag(Depth_m)<0)|is.na(lag(Depth_m)))%>%
  dplyr::summarize(buoy_freq = buoyancy.freq(Temp_C, Depth_m),
                   Depth_m = attr(buoyancy.freq(Temp_C, Depth_m), 'depths'))
buoy_max = buoyancy_raw%>%
  group_by(Date)%>%
  dplyr::summarize(max = max(buoy_freq))
buoy_max$Date_22 = buoy_max$Date
year(buoy_max$Date_22)=2022

write.csv(buoy_max, "../Processed_data/BF.csv", row.names = F)
```


Step 6: Calculate volume-weighted water temperature
```{r}
#Calculate volume-weighted water temperature each day
#First, need to add surface measurements where they are missing
surf = catwalk_format%>%
  group_by(Date)%>%
  dplyr::summarize(Reading = Reading[which.min(Sensor_depth)],
            Sensor_depth = 0,
            Reservoir_depth = unique(Reservoir_depth))
catwalk_format_with_zero = catwalk_format%>%
  full_join(surf)%>%
  arrange(Date,Sensor_depth)
#Then, run the calculations
vw_temp = c()
for(date in unique(catwalk_format_with_zero$Date)){
  temps = catwalk_format_with_zero%>%
    filter(Date == date)
  baths = flexible_bathy%>%
    filter(Date==date)
  vw_temp = c(vw_temp, layer.temperature(top = min(baths$Depth_m), bottom = max(baths$Depth_m), wtr = temps$Reading, depths = temps$Sensor_depth, bthA = baths$SA_m2, bthD = baths$Depth_m))
}
vw_temp_df = data.frame(Date = unique(catwalk_format$Date), vw_temp = vw_temp)%>%
  mutate(Year = year(Date))
vw_temp_df$Date_22 = vw_temp_df$Date
year(vw_temp_df$Date_22)=2022

write.csv(vw_temp, "../Processed_data/VW_temp.csv", row.names = F)
```


Step 7: Calculate thermocline depths
```{r}
flexible_bathy = cat_depths%>%
  mutate(Reservoir = "BVR")%>%
  full_join(bathy)%>%
  group_by(Date)%>%
  mutate(Depth_m = Depth_m - (max(Depth_m)-unique(WaterLevel)) + 1)%>%
  filter(Depth_m>=-1)

thermo_raw = catwalk_format%>%
  filter(!is.na(Reading))%>%
  mutate(Depth_simple = floor(Sensor_depth))%>%
  group_by(Date,Depth_simple)%>%
  dplyr::summarize(Temp_C = mean(Reading),
            Depth_m = mean(Sensor_depth),
            Reservoir_depth = mean(Reservoir_depth))%>%
  arrange(Date,Depth_m)%>%
  group_by(Date)%>%
  filter((Depth_m-lag(Depth_m)>.4)|(Depth_m-lag(Depth_m)<0)|is.na(lag(Depth_m)))%>%
  dplyr::summarize(thermo = thermo.depth(Temp_C, Depth_m),
            Reservoir_depth = mean(Reservoir_depth))

write.csv(thermo_raw,"../Processed_data/thermocline_depth.csv", row.names = F)
```


Step 8: Calculate volume of each layer of the reservoir
```{r}
meta_ratio = flexible_bathy%>%
  left_join(thermo_raw)%>%
  filter(!is.na(thermo),thermo>0)%>%
  group_by(Date)%>%
  rename(Volume_L = Volume_layer_L)%>%
  dplyr::summarize(depth_above_thermo = max(Depth_m[Depth_m<(thermo)]),
            thermo_and_just_below = sum(Volume_L[Depth_m<=depth_above_thermo]), #1m layers
            correction_thermo = unique(1-(thermo-depth_above_thermo))*Volume_L[Depth_m==depth_above_thermo],
            depth_above_water = unique(Depth_m[Depth_m<0]),
            correction_surf = abs(unique(depth_above_water))*Volume_L[Depth_m==depth_above_water],
            epi_volume = unique(thermo_and_just_below-correction_thermo-correction_surf),
            hypo_and_above = sum(Volume_L[Depth_m>=depth_above_thermo]),
            correction_hypo = unique(thermo-depth_above_thermo)*Volume_L[Depth_m==depth_above_thermo],
            hypo_volume = unique(hypo_and_above-correction_hypo),
            epi_hypo_ratio = unique(epi_volume/hypo_volume)
  )

meta_ratio$Date_22 = meta_ratio$Date
year(meta_ratio$Date_22) = 2022

write.csv(meta_ratio, "../Processed_data/meta_ratio.csv", row.names = F)
```