---
title: "Time series plots"
author: "Abby Lewis and Adrienne Breef-Pilz"
date: "2022-09-08"
output: html_document
---

This file creates three summary time series plots that are in the main manuscript (Figure 5-7).

To do so, this file loads physical, chemical, and biological data from other scripts. 

TABLE OF CONTENTS:
Code chunk 1: Load packages, set plot specifications
Code chunk 2: Load data
Code chunk 3: Compile Stats for Figure 5
Code chunk 4: Make line plot of nutrients, phytos, DOC, and CO2 (Figure 5)
Code chunk 5: Compile stats for Figure 6
Code chunk 6: Make line plot of turbidity, euphotic zone, peak width, and peak depth (Figure 6)
Code chunk 7: Compile stats for Figure 7
Code chunk 8: Make line plot of metals data, DO, and thermocline depth (Figure 7

Code chunk 1: Load packages, set plot specifications
```{r Load packages and define drawdown dates, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(akima)
library(colorRamps)
library(ggh4x)
library(patchwork)
source("../R/plot_facet.R")

#Define drawdown dates
lines_rect = data.frame(closed = as.Date("2022-06-28"), opened = as.Date("2022-05-19"), Year = 2022, year_class = "2022")

#Set theme for plots
theme_set(
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=9),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none",
        legend.title=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.spacing.y = unit(0, 'cm'),
        legend.text=element_text(size=7)
        )
  )

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figs")){dir.create("../Figs")}
```


Code chunk 2: Load data
```{r Load and format chemistry data}
#From Chem.Rmd
chem_full <- read_csv("../Processed_data/Chem at 0.1m.csv")

# From physical metrics script
thermo_depth <- read_csv("../Processed_data/thermocline_depth.csv")%>%
  filter(year(Date)%in%c(2022,2021))

# From CTD data processing script
light <- read_csv("../Processed_data/attenuation_calc.csv")%>%
  filter(year(Date)%in%c(2021,2022))

ctd <- read_csv("../Processed_data/CTD at 0.1m.csv")%>%
  filter(year(Date)%in%c(2021,2022))

# From GHG script
ghg <- read_csv("../Processed_data/GHG at 0.1m.csv")%>%
  filter(year(Date)%in%c(2021,2022))

# From metals script
metals <- read_csv("../Processed_data/Metals at 0.1m.csv")%>%
  filter(year(Date)%in%c(2021,2022))

# From fluora script
fluora_surface <- read_csv("../Processed_data/Fluora_surface.csv")%>%
  filter(year(Date)%in%c(2021,2022))

cmax_depth <- read_csv("../Processed_data/FP_CmaxDepth.csv")%>%
  filter(year(Date)%in%c(2021,2022))

peak_width <- read_csv("../Processed_data/FP_PeakWidth.csv")%>%
  filter(year(Date)%in%c(2021,2022))
```


Code chunk 3: Compile Stats for Figure 5
```{r bloom effects stats}
#Combine all data
figure_5_data <- chem_full%>%
  full_join(fluora_surface)%>%
  full_join(ghg)%>%
  select(Date, TP_ugL, TN_ugL, PO4_ugL, DOC_mgL, 
         TotalConc_ugL, NH4_ugL, CO2_umolL, Bluegreens_ugL)%>%
  pivot_longer(cols = c(TP_ugL, TN_ugL,PO4_ugL, DOC_mgL,
                        TotalConc_ugL, CO2_umolL, NH4_ugL, Bluegreens_ugL))%>% 
  filter(!is.na(value),
         month(Date)>4,
         month(Date)<9)%>%
  mutate(Year = year(Date),
         color = name,
         name = ifelse(name%in%c("TotalConc_ugL","Bluegreens_ugL"),"Phyto",name), 
         name = ifelse(name%in%c("TP_ugL","PO4_ugL"),"Phosphorus",name), 
         name = ifelse(name%in%c("TN_ugL","NO3NO2_ugL","NH4_ugL"),"Nitrogen",name))%>%
  mutate(name = factor(name, levels=c("Phosphorus","Nitrogen","Phyto","DOC_mgL","CO2_umolL")),
         color = factor(color, 
                        levels=c("TP_ugL","PO4_ugL","TN_ugL","NH4_ugL",
                                 "TotalConc_ugL","Bluegreens_ugL","DOC_mgL","CO2_umolL"),
                        labels=c("Total P","Phosphate","Total N","Ammonium",
                                 "Total phytoplankton", "Cyanobacteria", "DOC", "CO2")))

#Calculate maxes
maxes <- figure_5_data%>%
  group_by(name,color, Year)%>%
  summarize(max_date = Date[which.max(value)],
            min_date = Date[which.min(value)])%>%
  filter(Year=="2022")

maxes_sum <- figure_5_data%>%
  group_by(name,color, Year)%>%
  summarize(max = max(value))%>%
  group_by(name,color)%>%
  mutate(dif = max[Year == 2022] - max[Year == 2021],
         pct = (max[Year == 2022] - max[Year == 2021]) / max[Year == 2021] * 100)

#Stats
nut_maxes <- maxes %>%
  mutate(max_date=as.Date(max_date, origin="1970-01-01"))

nut_maxes$max_date[which(nut_maxes$color == 'DOC')] - 
  nut_maxes$max_date[which(nut_maxes$color == 'Total phytoplankton')] 

nut_maxes$max_date[which(nut_maxes$color == 'CO2')] - 
  nut_maxes$max_date[which(nut_maxes$color == 'Total phytoplankton')] 

nut_maxes$max_date[which(nut_maxes$color == 'Ammonium')] - 
  nut_maxes$max_date[which(nut_maxes$color == 'Total phytoplankton')] 

```

Code chunk 4: Make line plot of nutrients, phytos, DOC, and CO2 (Figure 5)
```{r plots bloom effect}
# Phosphorus
p <- plot_facet(data = figure_5_data, 
                var_name = "Phosphorus", 
                xint_1 = 19170,
                xint_2 = 19163,
                color = c("black","black"),
                ylab = expression(atop(NA, atop(textstyle("Phosphorus"), 
                                                textstyle(paste("(",mu, "g/L)"))))),
                legend = T,
                legend.position = c(0.12,0.38),
                strip.placement = "outside", 
                strip.text.x = element_text(size = 12, face = "bold", vjust = 2),
                legend.justification = c(0,0))

# Nitrogen
n <- plot_facet(data = figure_5_data, 
                var_name = "Nitrogen", 
                xint_1 = 19170,
                xint_2 = 19191,
                color = c("magenta4", "magenta4"),
                ylab = expression(atop(NA, atop(textstyle("Nitrogen"), 
                                                textstyle(paste("(",mu, "g/L)"))))),
                legend = T,
                legend.position = c(0.12, 0.38),
                legend.justification = c(0, 0))

# Phytoplankton
phyto <-plot_facet(data = figure_5_data, 
                   var_name = "Phyto", 
                   xint_1 = 19170,
                   xint_2 = 19170,
                   color = c("darkgreen", "darkgreen"),
                   ylab = expression(atop(NA, atop(textstyle("Phytoplankton"), 
                                                   textstyle(paste("(",mu, "g/L)"))))),
                   legend = T,
                   legend.position = c(0.12, 0.38),
                   legend.justification = c(0, 0))

# DOC
doc <- plot_facet(data = figure_5_data, 
                  var_name = "DOC_mgL", 
                  xint_1 = 19178,
                  color = "#8E4412",
                  ylab = expression(atop(NA, atop(textstyle("DOC"), 
                                                  textstyle(paste("(mg/L)")))))
                  )

# CO2
co2 <- plot_facet(data = figure_5_data, 
                var_name = "CO2_umolL", 
                xint_1 = 19184,
                color = "blue",
                ylab = expression(atop(NA, atop(textstyle("CO"[2]), 
                                                textstyle(paste("(", mu, "mol/L)"))))),
                axis.text.x = element_text(vjust = -1), 
                axis.ticks = element_line(colour = "black"),
               )

figure_5 <- wrap_plots(p, n, phyto, doc, co2, nrow=5)

ggsave("../Figs/Figure_5.jpeg", figure_5, dpi = 300, width = 6, height = 5, bg = "white")
```


Code chunk 5: Compile stats for Figure 6
```{r phytos stats}
figure_6_data <- fluora_surface%>%
  full_join(light)%>%
  full_join(ctd%>%select(-Year,-Depth_m))%>%
  full_join(cmax_depth)%>%
  full_join(peak_width)%>%
  select(Date, Turbidity_NTU, Zeu, CmaxDepth_TotalConc_ugL, 
         CmaxDepth_Bluegreens_ugL, PeakWidth_TotalConc_m, PeakWidth_Bluegreens_m)%>%
  pivot_longer(cols = c(Turbidity_NTU, Zeu, CmaxDepth_TotalConc_ugL, 
                        CmaxDepth_Bluegreens_ugL, PeakWidth_TotalConc_m, PeakWidth_Bluegreens_m))%>%
  filter(!is.na(value),
         month(Date)>4,
         month(Date)<9)%>%
  mutate(Year = year(Date),
         color = ifelse(name%in%c("CmaxDepth_TotalConc_ugL", "PeakWidth_TotalConc_m"),
                        "TotalConc_ugL",
                        ifelse(name%in%c("PeakWidth_Bluegreens_m", "CmaxDepth_Bluegreens_ugL"),
                               "Bluegreens_ugL",name)),
         name = ifelse(name%in%c("CmaxDepth_TotalConc_ugL", "CmaxDepth_Bluegreens_ugL"),
                       "Cmax_depth", name),
         name = ifelse(name%in%c("PeakWidth_TotalConc_m", "PeakWidth_Bluegreens_m"),
                       "PeakWidth",name),
         color = factor(color, 
                        levels = c("Zeu", "TotalConc_ugL", "Bluegreens_ugL", "Turbidity_NTU"),
                        labels = c("Zeu", "Total phytoplankton", "Cyanobacteria", "Turbidity_NTU"))
         )

#Calculate maxes
maxes <- figure_6_data%>%
  group_by(name,color, Year)%>%
  summarize(max_date = Date[which.max(value)],
            min_date = Date[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("Zeu", "Cmax_depth", "PeakWidth"),min_date,max_date))

#Stats
figure_6_data%>%
  group_by(name,color, Year)%>%
  summarize(max_date = Date[which.max(value)],
            max = max(value),
            min_date = Date[which.min(value)],
            min = min(value))%>%
  mutate(max_date = ifelse(name %in% c("Zeu", "Cmax_depth", "PeakWidth"),min_date,max_date),
         max = ifelse(name %in% c("Zeu", "Cmax_depth", "PeakWidth"),min,max))%>%
  group_by(name,color)%>%
  mutate(pct = (max[Year==2022]-max[Year==2021])/max[Year==2021]*100)
```

Code chunk 6: Make line plot of turbidity, euphotic zone, peak width, and peak depth (Figure 6)
```{r phytos with legend plot}
#PLOT
# plot each one individual to plots and then merge them at the end
turb <- plot_facet(data = figure_6_data, 
                   var_name = "Turbidity_NTU", 
                   xint_1 = 19178,
                   color = "black",
                   ylab = expression(atop(NA, atop(textstyle("Turbidity"), 
                                                   textstyle("(NTU)")))),
                   strip.placement = "outside", 
                   strip.text.x = element_text(size = 12, face = "bold", vjust = 2))

# Plots Euphotic Zone
zeu <- plot_facet(data = figure_6_data, 
                  var_name = "Zeu", 
                  xint_1 = 19178,
                  color = "goldenrod",
                  reverse = T,
                  ylab = expression(atop(NA, atop(textstyle("Euphotic zone"), 
                                                  textstyle("depth (m)")))))

# Cmax
cmax <- plot_facet(data = figure_6_data, 
                   var_name = "Cmax_depth", 
                   xint_1 = 19163,
                   xint_2 = 19178,
                   reverse = T,
                   color = c("darkgreen","darkgreen"),
                   ylab = expression(atop(NA, atop(textstyle("C"["max"]), 
                                                   textstyle("depth (m)")))),
                   legend = T,
                   legend.position = c(0.20,0.66))

# Peak width
width <- plot_facet(data = figure_6_data, 
                    var_name = "PeakWidth", 
                    xint_1 = 19156,
                    xint_2 = 19163,
                    color = c("darkgreen","darkgreen"),
                    ylab = expression(atop(NA, atop(textstyle("Peak width"), 
                                                    textstyle("(m)")))),
                    legend = T,
                    legend.position = c(0.30,0.77),
                    axis.text.x = element_text(vjust = -1), 
                    axis.ticks = element_line(colour = "black"))

figure_6 <- wrap_plots(turb, zeu, cmax, width, nrow=4)

ggsave("../Figs/Figure_6.jpeg", figure_6, dpi = 300, width = 6, height = 4, bg = "white")
```


Code chunk 7: Compile stats for Figure 7
```{r Stats for Oxygen figure}
#combine relevant data
figure_7_data <- thermo_depth%>%
  full_join(ctd %>% select(-Year,-Depth_m))%>%
  full_join(metals)%>%
  select(Date, TFe_mgL, TMn_mgL, SFe_mgL, SMn_mgL, DOsat_percent,thermo)%>%
  pivot_longer(cols = c(TFe_mgL, TMn_mgL, SFe_mgL, SMn_mgL, DOsat_percent, thermo))%>% #Pivot for facet wrap
  filter(!is.na(value),
         month(Date) > 4,
         month(Date) < 9)%>%
  mutate(Year = year(Date),
         color = name,
         name = ifelse(name %in% c("SFe_mgL", "TFe_mgL"), "Fe_mgL", name),
         name = ifelse(name %in% c("SMn_mgL", "TMn_mgL"), "Mn_mgL", name),
         color = factor(color, 
                        levels = c("thermo", "DOsat_percent", "TFe_mgL", "TMn_mgL", "SFe_mgL", "SMn_mgL"),
                        labels = c("thermo", "DOsat_percent", "Total Fe", "Total Mn", "Soluble Fe", "Soluble Mn")))

#Calculate maximum values
maxes <- figure_7_data%>%
  group_by(name, color, Year)%>%
  summarize(max_date = Date[which.max(value)],
            min_date = Date[which.min(value)])%>%
  filter(Year=="2022")%>%
  mutate(max_date = ifelse(name %in% c("thermo", "DOsat_percent"), 
                           min_date, 
                           max_date))

maxes_sum <- figure_7_data%>%
  group_by(name, color, Year)%>%
  summarize(max = max(value),
            max_date = Date[which.max(value)],
            min_date = Date[which.min(value)])%>%
  group_by(name, color)%>%
  mutate(dif = max[Year==2022] - max[Year==2021],
         pct = (max[Year==2022] - max[Year==2021]) / max[Year==2021] * 100,
         date_dif = max_date[Year == 2022] - (max_date[Year == 2021] + 365))

mins_sum <- figure_7_data%>%
  group_by(name, color, Year)%>%
  summarize(min = min(value),
            min_date = Date[which.min(value)],
            min_date = Date[which.min(value)])%>%
  group_by(name,color)%>%
  mutate(dif = min[Year == 2022] - min[Year == 2021],
         pct = (min[Year == 2022] - min[Year == 2021]) / min[Year == 2021] * 100,
         date_dif = min_date[Year == 2022]-(min_date[Year == 2021] + 365))

#Stats
figure_7_data%>%
  group_by(name, color, Year)%>%
  summarize(max_date = Date[which.max(value)],
            max = max(value),
            min_date = Date[which.min(value)],
            min = min(value))%>%
  mutate(max_date = ifelse(name %in% c("thermo", "DOsat_percent"),
                           min_date,
                           max_date))
```


Code chunk 8: Make line plot of metals data, DO, and thermocline depth (Figure 7)
```{r Plot for oxygen Figure}
thermo <- plot_facet(data = figure_7_data, 
                     var_name = "thermo", 
                     xint_1 = 19177, 
                     color = "darkblue",
                     ylab = "Thermocline \ndepth (m)",
                     reverse = T,
                     strip.placement = "outside", 
                     strip.text.x = element_text(size = 12, face = "bold", vjust = 2))

do_sat <- plot_facet(data = figure_7_data, 
                     var_name = "DOsat_percent", 
                     xint_1 = 19184,
                     color = "cyan4",
                     ylab = "Surface DO\n(% saturation)")

# Iron
fe <- plot_facet(data = figure_7_data, 
                 var_name = "Fe_mgL", 
                 xint_1 = 19184,
                 xint_2 = 19178,
                 color = c("lightsalmon","lightsalmon"),
                 ylab = "Fe (mg/L)",
                 legend = T,
                 legend.position = c(0.18,0.66))

# Mn
mn <- plot_facet(data = figure_7_data, 
                 var_name = "Mn_mgL", 
                 xint_1 = 19184,
                 xint_2 = 19170,
                 color = c("coral4","coral4"),
                 ylab = "Mn (mg/L)",
                 legend = T,
                 axis.text.x = element_text(vjust = -1), 
                 axis.ticks = element_line(colour = "black"), 
                 legend.position = c(0.18,0.66))

figure_7 <- wrap_plots(thermo, do_sat, fe, mn, nrow=4)

ggsave(file="../Figs/Figure_7.jpeg",figure_7, dpi = 300, width = 6, height = 4, bg = "white")
 
```
