---
title: "Physics"
author: "Abby Lewis"
date: "2022-08-30"
output: html_document
---

This file uses thermistor and meteorological data to create lake physics plots (Figures 2-4).

Step 1: Load packages and data, and identify drawdown dates for figures
Step 2: Plot lake physics metrics (Schmidt, BF, VW temp, surface temp)
Step 3: Plot meteorology
Step 4: Plot volume, divided by layer
Step 5: Create a heatmap of water temperature over time
Step 6: Create a heatmap of DO over time
Step 7: Stats for paper


Step 1: Load packages and data, and identify drawdown dates for figures
```{r setup, include=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(akima)
library(colorRamps)
library(ggpubr)
library(cowplot)
source("../R/thermo.depth.density.R")

# Load data
schmidt_df <- read_csv("../Processed_data/Schmidt.csv")
buoy_max <- read_csv("../Processed_data/BF.csv")
vw_temp_df <- read_csv("../Processed_data/VW_temp.csv")
cat_long <- read_csv("../Processed_data/BVR_longoutput.csv")
meta_ratio <- read_csv("../Processed_data/meta_ratio.csv")
weather_sum  <- read_csv("../Processed_data/Weather_sum.csv")%>%
  mutate(Date_22 = DOY + as.Date("2021-12-31"))
thermo_raw <- read_csv("../Processed_data/thermocline_depth.csv")
df.final <- read_csv("../Processed_data/CTD trimmed.csv")
met_drawdown <- read_csv("../Processed_data/Weather_drawdown.csv")
sa <- read_csv("../Processed_data/Dynamic_surface_area.csv")

# Set lines
lines <- data.frame(Dates = as.Date(c("2022-06-28", "2022-05-19")), 
                   desc = c("closed","opened"), 
                   type = c("close","open"),
                   status = c("init","init"), 
                   Year = 2022,
                   year_class = "2022")
# Re-format for rectangle instead of discrete lines
lines_rect <- lines%>%
  select(-type)%>%
  pivot_wider(names_from = desc, values_from = Dates)

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figs")){dir.create("../Figs")}
```


Step 2: Plot lake physics metrics (Schmidt, BF, VW temp, surface temp)
```{r}
schmidt_fig <- schmidt_df%>%
  filter(Year %in% c(2021,2022),
         !is.na(Schmidt))%>%
  mutate(Year = as.factor(Year),
         Date_22 = DOY + as.Date("2021-12-31"))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22,y = Schmidt, fill = Year, color = Year), 
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  ylab(expression(paste("Schmidt stability ", "(J/",m^2, ")")))+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank(),
        legend.position = "none")
schmidt_fig

buoy_fig <- buoy_max%>%
  mutate(Year = as.factor(year(Date)),
         Date_22 = DOY + as.Date("2021-12-31"))%>%
  filter(Year %in% c(2021,2022))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22, y= max, color = Year, fill = Year),
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  theme_bw()+
  ylab("Maximum buoyancy\nfrequency (1/s)")+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank(),
        legend.position = "none")
buoy_fig

water <- vw_temp_df%>%
  filter(Year %in% c(2021,2022),
         !is.na(vw_temp))%>%
  mutate(Year = year(Date),
         Year = as.factor(Year),
         Date_22 = DOY + as.Date("2021-12-31"))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22, y = vw_temp, fill = Year, color = Year),
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  ylab("Volume-weighted\nwater temperature (ºC)")+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank(),
        legend.position = "none")
water

surf <- cat_long%>%
  mutate(Date = as.Date(DateTime),
         Year = year(Date))%>%
  filter(Sensor_depth>0)%>%
  group_by(Date, Year)%>%
  dplyr::summarize(Reading = Reading[which.min(Sensor_depth)],
            Sensor_depth = mean(Reservoir_depth))%>%
  mutate(Year = year(Date),
         DOY = yday(Date),
         #Making a date column where year is the same to standardize x-axes
         Date_22 = DOY + as.Date("2021-12-31"))%>% 
  filter(Year %in% c(2021,2022))

surf_plot_save <- surf%>%
  mutate(Year = as.factor(Year))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22, y = Reading, fill = Year, color = Year), 
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  ylab("Surface water temperature (ºC)")+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank())+
  theme(legend.position = "none")
surf_plot_save
```


Step 3: Plot meteorology
```{r}
temp <- weather_sum%>%
  mutate(Year = as.factor(Year),
         Date_22 = DOY + as.Date("2021-12-31"))%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22, y = AirTemp_C_Average, fill = Year, color = Year), 
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  ylab("Air temperature (ºC)")+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank())

legend <- get_legend(temp)
jpeg("../Figs/Figure_4.jpg",res = 300, units = "in",width = 8, height = 6)
plot_grid(temp+theme(legend.position = "none"),
          surf_plot_save,
          water,
          buoy_fig,
          schmidt_fig,
          legend, 
          align = "hv",
          labels = list("a","b","c","d","e",""),
          nrow=2,ncol=3,axis = "rltb")
dev.off()

## SI meteorology plots: rain and wind

# Rain
cum_rain = weather_sum%>%
  group_by(Year)%>%
  mutate(cum_rain_mm = cumsum(Rain_Total_mm),
         Year = as.factor(Year),
         Date_22 = DOY + as.Date("2021-12-31"))
rain_cum = cum_rain%>%
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_line(aes(x = Date_22, y = cum_rain_mm, color = Year),size = 1.5)+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  scale_linetype_manual(values = c("solid","dashed"))+
  ylab("Cumulative rain (mm)")+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank())
# Wind
wind = weather_sum%>%
  mutate(Year = as.factor(Year),
         Wind = WindSpeed_Average_m_s,
         #Making a date column where year is the same to standardize x-axes
         Date_22 = DOY + as.Date("2021-12-31"))%>% 
  ggplot()+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "grey40", data = lines_rect, alpha = 0.2)+
  geom_point(aes(x = Date_22, y = Wind, fill = Year, color = Year), 
             shape = 21, stroke = .5, size = 1.5)+
  scale_fill_manual(values = c("black","#00F7FF"))+
  scale_color_manual(values = c("black","#0800FA"))+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  ylab("Wind speed (m/s)")+
  theme_bw()+
  scale_x_date(date_breaks = "2 month", date_labels =  "%b")+
  theme(axis.title.x = element_blank())

jpeg("../Figs/SI_met.jpg",res = 300, units = "in",width = 6, height = 4)
ggarrange(rain_cum, wind, legend = "bottom", labels = "auto")
dev.off()
```


Step 4: Plot volume, divided by layer
```{r}
jpeg("../Figs/Figure_2c.jpeg",width = 6.5, height = 3, units = "in", res = 300)
meta_ratio %>%
  mutate(Year = as.factor(year(Date)),
         DOY = yday(Date),
         Date_22 = DOY + as.Date("2021-12-31"))%>%
  filter(!Year==2020,
         month(Date)>=5,
         month(Date)<10)%>%
  dplyr::select(Date_22, Year, epi_volume,hypo_volume)%>%
  pivot_longer(cols = c(epi_volume,hypo_volume))%>%
  mutate(name = ifelse(name=="epi_volume",
                       "Epilimnetic volume",
                       "Hypolimnetic volume"),
         name = factor(name,levels = c("Epilimnetic volume",
                                       "Hypolimnetic volume")))%>%
  ggplot()+
  scale_fill_manual(values = c("#77c7fc","#104466"))+
  geom_area(aes(x = Date_22, y = value/1000, fill = name))+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "black", data = lines_rect, alpha = 0.15)+
  geom_vline(data=lines,aes(xintercept = Dates, lty = status), show.legend = F)+
  scale_linetype_manual(values = c("solid","dashed"))+
  facet_grid(rows = ~Year, scales = "free_x", space = "free_x")+
  theme_bw()+
  ylab(bquote('Volume'~(m^3)))+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
dev.off()
```


Step 5: Create a heatmap of water temperature over time
```{r}
var <- c("Reading") #Temp_C

filt_no_surf <- cat_long%>%
  mutate(Date = as.Date(DateTime),
         Year = year(Date),
         Height = Reservoir_depth-Sensor_depth,
         Rounded_depth_tenth = round(Height,1))%>%
  filter(!is.na(get(var)),
         Reservoir == "BVR")%>%
  ungroup()%>%
  dplyr::select(Date, Rounded_depth_tenth,Reading, Year)%>%
  group_by(Year, Date, Rounded_depth_tenth)%>%
  dplyr::summarize(Reading=mean(Reading, na.rm = T))%>%
  dplyr::rename(Sensor_depth = Rounded_depth_tenth)

surf <- cat_long%>%
  filter(Sensor_depth >= 0) %>%
  mutate(Date = as.Date(DateTime),
         Year = year(Date))%>%
  group_by(Date, Year)%>%
  dplyr::summarize(Reading = Reading[which.min(Sensor_depth)],
            Sensor_depth = mean(Reservoir_depth))

bot <- cat_long%>%
  mutate(Date = as.Date(DateTime),
         Year = year(Date))%>%
  group_by(Date, Year)%>%
  dplyr::summarize(Reading = Reading[which.max(Sensor_depth)],
            Sensor_depth = 0)
  
filt <- filt_no_surf%>%
  full_join(surf)%>%
  full_join(bot)

depth_sum <- cat_long%>%
  mutate(Date = as.Date(DateTime),
         Year = year(Date))%>%
  filter(Reservoir == "BVR")%>%
  ungroup()%>%
  dplyr::select(Date, Reservoir_depth,Year)%>%
  group_by(Year, Date)%>%
  dplyr::summarize(Reservoir_depth=mean(Reservoir_depth, na.rm = T))%>%
  ungroup()%>%
  filter(!Year == 2020)%>%
  mutate(max = max(Reservoir_depth))

Years <- c(2021,2022)
interp <- data.frame(x = NA, y = NA, z = NA, Site = NA)
for(year in Years){
    bvr <- filt%>%
      filter(Year == year)%>%
      ungroup()%>%
      dplyr::mutate(random = sample(1:n(),n()))%>%
      arrange(random)%>%
      filter(!is.na(Reading))
    
    if(nrow(bvr)>0){
      bvr_interp <- interp2xyz(interp(bvr$Date,bvr$Sensor_depth,bvr[[var]],
                                      xo = seq(min(bvr$Date), max(bvr$Date), 1),
                                      yo = seq(min(bvr$Sensor_depth), max(bvr$Sensor_depth), by = .05), 
                                      duplicate = "mean"),
                               data.frame = T)
      interp = interp%>%
        full_join(bvr_interp)
    }
    
}

sensor_depth_plot <- filt_no_surf%>%
  filter(Year%in% c(2021,2022))%>%
  group_by(Date)%>%
  filter(length(unique(Sensor_depth))<=11)%>%
  ungroup()%>%
  select(Year,Sensor_depth)%>%
  unique()%>%
  mutate(Date=as.Date(paste0(Year,"-01-01")))

custom_heatmap_palette <- c("#313695", "#4575b4", "#74add1",
                            "#abd9e9", "#e0f3f8", "#ffffbf", 
                            "#fee090","#fdae61", "#f46d43", 
                            "#d73027", "#a50026")

#Heatmap
jpeg(paste0("../Figs/BVR_catwalk_heatmap_2022.jpeg"),width = 8, height = 4, units = "in", res = 300)
water_temp <- interp%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=x, y=y,fill = z))+
  geom_ribbon(data = depth_sum, 
              aes(x = Date, ymin= Reservoir_depth, ymax = max), 
              fill = "white", color = "black",lwd=.3)+
  geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
            fill = "black", data = lines_rect, alpha = 0.15)+
  geom_line(data = thermo_raw%>%
              mutate(Year=year(Date))%>%
              filter(Year%in%c(2021,2022)), 
            aes(x = Date, y = Reservoir_depth-thermo, lty = "Thermocline"), 
            lwd = .2)+
  geom_hline(aes(yintercept = 6.6, lty = "Outtake pipe"), alpha = 0.3)+
  labs(y = "Height above sediments (m)")+
  scale_x_date(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank())+
  facet_grid(cols = vars(Year),
             scales = "free",
             space = "free"
             )+
  labs(color = "")+
  geom_point(data = sensor_depth_plot, 
             aes(x = Date, y = Sensor_depth), 
             shape = 23, fill = "black", color = "white", size  =2.5)+
  geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  scale_linetype_manual(values = c("solid","84"),
                        limits = c("Thermocline","Outtake pipe"),  
                        name= NULL)+
  scale_fill_gradientn(name = "Water temperature\n(°C)",
                       limits = c(-1, 30),
                       colours = custom_heatmap_palette)
print(water_temp)
dev.off()
```


Step 6: Create a heatmap of DO over time
```{r}
sensor_depth <- cat_long%>%
  mutate(Year = year(DateTime),
         Date = as.Date(DateTime),
         max = max(Reservoir_depth))%>%
  filter(Reservoir_depth > 5,
         Year %in% c(2021,2022))%>%
  group_by(Date,Year)%>%
  summarize(Reservoir_depth = max(Reservoir_depth),
            max = unique(max))%>%
  select(Date,Reservoir_depth,max,Year)

ctd_sensor <- df.final%>%
  left_join(sensor_depth%>%select(-max))%>%
  group_by(Date)%>%
  mutate(max_ctd = max(Depth_m),
         dif = Reservoir_depth-max_ctd)%>%
  mutate(Depth_m = Depth_m_sed + dif)%>%
  filter(!is.na(Depth_m))

bottom <- df.final%>%
  left_join(sensor_depth%>%select(-max))%>%
  filter(!is.na(Reservoir_depth))%>%
  group_by(Date)%>%
  filter(Depth_m_sed == min(Depth_m_sed))%>%
  mutate(Depth_m = 0)
  
ctd_adj_depth <- ctd_sensor%>%
  full_join(bottom)

vars <- c("DO_mgL")
names <- c("Dissolved oxygen\n(mg/L)")
Years <- 2021:2022
Sites <- c(50)

for(i in 1:length(vars)){
  var <- vars[i]
  interp <- data.frame(x = NA, y = NA, z = NA, Site = NA)
  for(site in Sites){
    for(year in Years){
      bvr <- ctd_adj_depth%>%
        mutate(Year = year(Date),
               Date = as.Date(Date))%>%
        filter(Reservoir == "BVR",
               !is.na(get(var)),
               !is.na(Depth_m),
               Site %in% Sites,
               Year %in% Years)%>%
        group_by(Site)%>%
        mutate(max = max(Depth_m))
      bvr_site <- bvr%>%
        filter(Site == site,
               Year == year)
      if(nrow(bvr_site)>0){
        bvr_interp <- interp2xyz(interp(bvr_site$Date,bvr_site$Depth_m,bvr_site[[var]],
                                        xo = seq(min(bvr_site$Date), max(bvr_site$Date), .1),
                                        yo = seq(min(bvr_site$Depth_m), max(bvr_site$Depth_m), 
                                                 by = .01), 
                                        duplicate = "mean"),data.frame = T)
        interp <- interp%>%
          full_join(bvr_interp%>%mutate(Site=site))
      }
    }
  }
  interp <- interp%>%
    filter(!is.na(Site))
  
  bvr_maxes <- bvr%>%
    group_by(Site)%>%
    summarize(max = unique(max))
  
  #Heatmap
  jpeg(paste0("../Figs/BVR_",var,"_heatmap_2022.jpeg"),width = 8, height = 4, units = "in", res = 300)
  ctd_plot <- interp%>%
    mutate(x = as.Date(x, origin = "1970-01-01"),
           Year = year(x))%>%
    ggplot()+
    geom_raster(aes(x=x,fill = z, y=y))+
    scale_y_continuous(expand = c(0,0),limits = c(0, NA))+
    labs(y = "Height above sediments (m)")+
    scale_fill_gradientn(colours = if(var%in%c("DO_mgL","DO_pSat","ORP_mV")){rev(custom_heatmap_palette)}else{custom_heatmap_palette}, 
      na.value="white", name = names[i])+
    geom_ribbon(data = sensor_depth, 
                aes(x = Date, ymin= Reservoir_depth, ymax = max), 
                fill = "white",color = "black",lwd=.2)+
    geom_point(data = bvr, aes(x = Date, y = max), 
               shape = 25, fill = "black", color = "white", size  =2)+
    geom_rect(aes(xmin = opened,xmax = closed,ymin=-Inf,ymax=Inf),
              fill = "black", data = lines_rect, alpha = 0.15)+
    geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
    geom_hline(aes(yintercept = 6.6, lty = "Outtake pipe"), 
               alpha = 0.3, lty = "84")+
    scale_x_date(expand = c(0,0))+
    theme(panel.border = element_rect(fill = NA),
          panel.background = element_rect(fill = "white"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title.x=element_blank())+
    facet_grid(cols = vars(Year),
               scales = "free",
               space = "free_y")
  print(ctd_plot)
  dev.off()
}

jpeg(paste0("../Figs/Figure_3.jpeg"),width = 6, height = 6, units = "in", res = 300)
ggarrange(water_temp,ctd_plot, nrow=2, align = "hv", labels = "auto")
dev.off()
```


Step 7: Stats for paper
```{r}
#### Drawdown extent

# How much did volume change as a result of drawdown?
vol_change <- meta_ratio %>%
  mutate(Year = year(Date),
         DOY = yday(Date),
         timepoint = ifelse(DOY == yday("2022-05-19"), "start",
                            ifelse(DOY == yday("2022-06-28"), "end", 
                                   NA))) %>%
  filter(!is.na(timepoint)) %>%
  select(Year, timepoint, total_volume, epi_volume, hypo_volume) %>%
  pivot_wider(names_from = timepoint, 
              values_from = c(total_volume, epi_volume, hypo_volume)) %>%
  mutate(total_change_L = total_volume_end - total_volume_start,
         total_change_pct = total_change_L/total_volume_start*100,
         epi_change_L = epi_volume_end - epi_volume_start,
         epi_change_pct = epi_change_L/epi_volume_start*100,
         hypo_change_L = hypo_volume_end - hypo_volume_start,
         hypo_change_pct = hypo_change_L/hypo_volume_start*100)

vol_change$epi_change_L/vol_change$total_change_L*100

# How much did surface area change as a result of drawdown?
sa %>%
  mutate(timepoint = ifelse(DOY == yday("2022-05-19"), "start",
                            ifelse(DOY == yday("2022-06-28"), "end", 
                                   NA))) %>%
  filter(!is.na(timepoint)) %>%
  select(Year, timepoint, linear_approx_SA) %>%
  pivot_wider(names_from = timepoint, 
              values_from = linear_approx_SA) %>%
  mutate(change_m2 = end - start,
         change_pct = change_m2/start*100)

# How much did depth change as a result of drawdown?
cat_long %>%
  mutate(Date = as.Date(DateTime))%>%
  group_by(Date)%>%
  summarize(Reservoir_depth = mean(Reservoir_depth, na.rm = T))%>%
  mutate(Year = year(Date),
         DOY = yday(Date),
         timepoint = ifelse(DOY == yday("2022-05-19"), "start",
                            ifelse(DOY == yday("2022-06-28"), "end", 
                                   NA))) %>%
  filter(!is.na(timepoint)) %>%
  select(Year, timepoint, Reservoir_depth) %>%
  pivot_wider(names_from = timepoint, 
              values_from = Reservoir_depth) %>%
  mutate(change_m = end - start,
         change_pct = change_m/start*100)


#### Temperature dynamics
#Turnover dates
turnover_dates <- surf%>%
  rename(surf_temp = Reading)%>%
  select(-Sensor_depth)%>%
  left_join(bot%>%rename(bot_temp=Reading))%>%
  mutate(Year=year(Date))%>%
  filter(month(Date)>7)%>%
  group_by(Year)%>%
  arrange(Date)%>%
  summarize(turnover=first(Date[surf_temp-bot_temp<1]))

#How much earlier was turnover?
yday(turnover_dates$turnover[which(turnover_dates$Year == 2021)]) - yday(turnover_dates$turnover[which(turnover_dates$Year == 2022)])

mean_temps <- vw_temp_df%>%
  filter(month(Date) %in% c(7))%>%
  group_by(Year) %>%
  dplyr::summarise(mean = mean(vw_temp, na.rm = T),
                   sd = sd(vw_temp, na.rm = T))

# change in vw_temp
mean_temps$mean[which(mean_temps$Year == 2021)] - 
  mean_temps$mean[which(mean_temps$Year == 2022)]

# change in surface temp
surf%>%
  filter(month(Date) %in% c(7))%>%
  group_by(Year) %>%
  dplyr::summarise(mean = mean(Reading, na.rm = T),
                   sd = sd(Reading, na.rm = T))

mean_buoy <- buoy_max%>%
  filter(month(Date)%in%c(7))%>%
  mutate(Year = year(Date))%>%
  group_by(Year)%>%
  dplyr::summarise(mean = mean(max, na.rm = T),
                   sd = sd(max, na.rm = T))

# change in buoyancy frequency
(mean_buoy$mean[which(mean_buoy$Year == 2022)] - mean_buoy$mean[which(mean_buoy$Year == 2021)]) / 
    mean_buoy$mean[which(mean_buoy$Year == 2021)]

mean_schmidt <- schmidt_df%>%
  filter(month(Date)%in%c(7))%>%
  group_by(Year)%>%
  dplyr::summarise(mean = mean(Schmidt, na.rm = T),
                   sd = sd(Schmidt, na.rm = T))

# change in schmidt stability
(mean_schmidt$mean[which(mean_schmidt$Year == 2021)] - mean_schmidt$mean[which(mean_schmidt$Year == 2022)]) / mean_schmidt$mean[which(mean_schmidt$Year == 2021)]


#### Meteorology

# Average wind speed 
avg_wind<- met_drawdown%>%
  group_by(Year)%>%
  summarise(mean_wind=mean(WindSpeed_Average_m_s, na.rm=T), 
            sd_wind=sd(WindSpeed_Average_m_s, na.rm=T ))
# Average air temperatures
avg_AT <- met_drawdown%>%
  group_by(Year) %>% 
  summarise(mean_air=mean(AirTemp_C_Average, na.rm=T), 
            sd_air=sd(AirTemp_C_Average, na.rm=T ))
# Average daily precipitation 
rain <- met_drawdown%>%
  mutate(Date = as.Date(DateTime))%>%
  group_by(Date, Year)%>%
  summarise(total_rain_day=sum(Rain_Total_mm))
avg_rain <- rain%>%
  group_by(Year)%>%
  summarise(mean_rain=mean(total_rain_day, na.rm=T), 
            sd_rain=sd(total_rain_day, na.rm=T))
rain_annual <- weather_sum%>%
  group_by(Date, Year)%>%
  summarise(total_rain_day=sum(Rain_Total_mm))
avg_rain_annual <- rain_annual%>%
  group_by(Year)%>%
  summarise(mean_rain=mean(total_rain_day, na.rm=T), 
            sd_rain=sd(total_rain_day, na.rm=T))
```