---
title: "Fluora"
author: "Abby Lewis and Mary Lofton"
date: "Last updated: 2023-07-25 ASL"
output: html_document
---

This file loads fluoroprobe data from EDI and creates one SI figure, as well as summary statistics

Step 1: Load packages and FP data
Step 2: Compare with photic zone depth and create figure for SI
Step 3: Calculate statistics for manuscript


Step 1: Load packages and FP data
```{r setup, include=FALSE}
#load packages
library(tidyverse)
library(lubridate)

flo_metrics <- read.csv("../Processed_data/Fluora_bvr.csv")%>%
  mutate(Date = as.Date(Date))

# Make a directory for Figures if there isn't one
if (!dir.exists("../Figs")){dir.create("../Figs")}
```


Step 2: Compare with photic zone depth and create figure for SI
```{r}
#read in light attenuation calculations and look @ Cmax relative to pz depth
pz <- read_csv("../Processed_data/attenuation_calc.csv") %>%
  mutate(Date = date(Date))

Cmax = read.csv("../Processed_data/FP_CmaxDepth.csv")
Cmax <- tibble(Cmax) %>%
  mutate(Date = date(Date))

cmpz <- left_join(Cmax, pz, by = "Date") %>%
  mutate(green_rel_pz = (Zeu-CmaxDepth_GreenAlgae_ugL)/Zeu,
         cyano_rel_pz = (Zeu-CmaxDepth_Bluegreens_ugL)/Zeu,
         brown_rel_pz = (Zeu-CmaxDepth_BrownAlgae_ugL)/Zeu,
         crypto_rel_pz = (Zeu-CmaxDepth_MixedAlgae_ugL)/Zeu,
         total_rel_pz = (Zeu-CmaxDepth_TotalConc_ugL)/Zeu,
         green_rel_pz_0.1 = (Zeu_0.1-CmaxDepth_GreenAlgae_ugL)/Zeu_0.1,
         cyano_rel_pz_0.1 = (Zeu_0.1-CmaxDepth_Bluegreens_ugL)/Zeu_0.1,
         brown_rel_pz_0.1 = (Zeu_0.1-CmaxDepth_BrownAlgae_ugL)/Zeu_0.1,
         crypto_rel_pz_0.1 = (Zeu_0.1-CmaxDepth_MixedAlgae_ugL)/Zeu_0.1,
         total_rel_pz_0.1 = (Zeu_0.1-CmaxDepth_TotalConc_ugL)/Zeu_0.1) %>%
  select(-(CmaxDepth_GreenAlgae_ugL:r2))%>%
  filter(!is.na(Zeu)) %>%
  mutate(Year = year(Date)) %>%
  gather(green_rel_pz:total_rel_pz_0.1, key = "spectral_group", value = "Cmax_depth_rel_pz") %>%
  mutate(pz_calc = ifelse(grepl("0.1",spectral_group),"0.1%","1%"))

#data wrangling to plot
Cmax_plot <- Cmax %>%
  mutate(Year = year(Date)) %>%
  gather(CmaxDepth_GreenAlgae_ugL:CmaxDepth_TotalConc_ugL, key = "spectral_group", value = "Cmax_depth")

cmpz_1 <- cmpz %>% filter(pz_calc == "1%")
cmpz_0.1 <- cmpz %>% filter(pz_calc == "0.1%")

#data wrangling to plot Zeu
pz_plot <- pz %>%
  mutate(Year = year(Date)) %>%
  filter(Year %in% c(2021,2022)) %>%
  select(-(I0:r2)) %>%
  gather(Zeu:Zeu_0.1, key = "perc_light", value = "pz_depth")

jpeg("../Figs/SI_Cmax_depth_pz_1.jpeg", res = 300, width = 6, height = 4, units = "in")
ggplot(data = subset(cmpz_1, cmpz_1$spectral_group != "crypto_rel_pz"), aes(x = spectral_group, y = Cmax_depth_rel_pz, fill = as.factor(Year)))+
  geom_boxplot()+
  geom_hline(yintercept = 0)+
  theme_bw()+
  ggtitle("Photic zone = 1% light; this is ONLY May-July")+
  xlab("Spectral group")+
  ylab("Cmax depth standardized to photic zone depth")+
  labs(fill = "Year")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_discrete(labels=c("green_rel_pz" = "Green algae", "cyano_rel_pz" = "Cyanobacteria",
                              "brown_rel_pz" = "Brown algae", "total_rel_pz" = "Total phyto."))
dev.off()

jpeg("../Figs/SI_Cmax_depth_pz_0.1.jpeg", res = 300, width = 6, height = 4, units = "in")
ggplot(data = subset(cmpz_0.1, cmpz_0.1$spectral_group != "crypto_rel_pz_0.1"), aes(x = spectral_group, y = Cmax_depth_rel_pz, fill = as.factor(Year)))+
  geom_boxplot()+
  geom_hline(yintercept = 0)+
  theme_bw()+
  ggtitle("Photic zone = 0.1% light; this is ONLY May-July")+
  xlab("Spectral group")+
  ylab("Cmax depth standardized to photic zone depth")+
  labs(fill = "Year")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_discrete(labels=c("green_rel_pz_0.1" = "Green algae", "cyano_rel_pz_0.1" = "Cyanobacteria",
                              "brown_rel_pz_0.1" = "Brown algae", "total_rel_pz_0.1" = "Total phyto."))
dev.off()
```


Step 3: Calculate statistics for manuscript
```{r}
# dominance of cyanobacteria during study period
dom <- flo_metrics %>%
  select(-Reservoir, -Site) %>%
  mutate(Date = date(DateTime)) %>%
  filter((Date >= "2021-05-01" & Date <= "2021-09-01") | (Date >= "2022-05-01" & Date <= "2022-09-01")) %>%
  group_by(Date) %>%
  summarize(GreenAlgae_ugL = mean(GreenAlgae_ugL, na.rm = TRUE),
            Bluegreens_ugL = mean(Bluegreens_ugL, na.rm = TRUE),
            BrownAlgae_ugL = mean(BrownAlgae_ugL, na.rm = TRUE),
            MixedAlgae_ugL = mean(MixedAlgae_ugL, na.rm = TRUE),
            TotalConc_ugL = mean(TotalConc_ugL, na.rm = TRUE)) %>%
  mutate(prop_ga = GreenAlgae_ugL/TotalConc_ugL,
         prop_bg = Bluegreens_ugL/TotalConc_ugL,
         prop_ba = BrownAlgae_ugL/TotalConc_ugL,
         prop_ma = MixedAlgae_ugL/TotalConc_ugL) %>%
  rowwise() %>%
  mutate(dom = max(c(prop_ga, prop_bg, prop_ba, prop_ma))) %>%
  rowwise() %>%
  mutate(dom_group = ifelse(dom == prop_ga, "GreenAlgae",
                            ifelse(dom == prop_bg, "Bluegreens",
                                   ifelse(dom == prop_ba, "BrownAlgae",
                                          ifelse(dom == prop_ma, "MixedAlgae","zombies")))))

dom %>%
  count(dom_group)
```